---
title: "Systematic CV-RCT Bayesian Pipeline"
subtitle: "EB + MCMC + Type M/S + S-values across a curated cardiovascular RCT cohort"
author: "Dr. James Brophy"
format:
  html:
    theme: [simple, custom.scss]
    toc: true
    toc-depth: 2
    code-fold: true
    embed-resources: true
execute:
  echo: true
  warning: false
  message: false
  cache: true
  freeze: auto
bibliography: bib.bib
csl: vancouver.csl
params:
  n_mcmc_chains: 2
  n_mcmc_iter: 1000
  n_eb_draws: 100000
---

```{r}
#| label: setup
#| include: true
#| cache: false
suppressPackageStartupMessages({
  library(dplyr)
  library(tidyr)
  library(tibble)
  library(readr)
  library(ggplot2)
  library(purrr)
})

# Shared Shiny/EB helpers (merged post-PR1) + pipeline-specific helpers.
source("R/priors.R")
source("R/eb_mixture.R")
source("R/pipeline.R")

set.seed(2026)

RESULTS_DIR <- "results"
dir.create(RESULTS_DIR, showWarnings = FALSE, recursive = TRUE)

# Tuned-down MCMC for fast end-to-end render. Bump chains/iter for production.
N_CHAINS <- params$n_mcmc_chains
N_ITER   <- params$n_mcmc_iter
N_EB     <- params$n_eb_draws

```

```{r}
#| label: stan-probe
#| include: false
#| cache: false
# Detect a Stan backend ONCE. We (a) check for cmdstanr/rstan, then (b) do
# a real micro-compile. Any compilation error (including broken C++
# toolchains that write to the subprocess stderr) is hidden because the
# whole chunk is rendered with `include: false` - output, messages and
# warnings are all suppressed before they can reach the HTML. On any
# failure we fall straight through to the analytic Normal-Normal path,
# which is mathematically identical to the brms normal-pseudo-likelihood
# model for this problem.
HAVE_STAN <- tryCatch({
  b <- tryCatch(choose_brms_backend(), error = function(e) NULL)
  if (is.null(b) || !nzchar(b) || !requireNamespace("brms", quietly = TRUE)) {
    FALSE
  } else {
    fit <- suppressMessages(suppressWarnings(tryCatch(
      brms::brm(
        formula = brms::bf(effect | se(se) ~ 1),
        data    = tibble::tibble(effect = 0, se = 1),
        family  = gaussian(),
        prior   = brms::prior_string("normal(0, 10)", class = "Intercept"),
        backend = b, chains = 1, iter = 200, refresh = 0, silent = 2, seed = 1
      ),
      error = function(e) NULL
    )))
    !is.null(fit)
  }
}, error = function(e) FALSE)
```

## Overview

This document implements a systematic, reproducible re-analysis of a curated
cohort of landmark cardiovascular RCTs. For each trial we compute:

- Observed log-HR, SE, z, p-value, and Shannon S-value.
- Type M / Type S errors via Gelman–Carlin retrodesign simulation.
- Empirical-Bayes shrinkage on the signal-to-noise-ratio (SNR) scale using the
  4-component van Zwet et al. 2025 mixture [@RN2].
- Five MCMC posteriors under different priors via `brms` with a normal
  pseudo-likelihood on (log-HR, SE_logHR):
  Weak, Skeptical, Enthusiastic, EB, Design-informed.

The last section reproduces the DECAF worked example from `eb2.qmd` through
this pipeline and sanity-checks the EB posterior HR.

Citation: van Zwet E, Więcek W, Gelman A. Meta-analysis with a single study.
*Stat Methods Med Res* 2025;34(12):2302-2312. doi:10.1177/09622802251380628.

### Headline findings

- Posterior `P(HR < 1)` is essentially 1 for all studies that crossed
  significance, but `P(HR ≤ MCID)` varies from ~0.05 (EMPA-REG) to ~1
  (PARADIGM-HF, RALES) under the EB prior — i.e., *clinically meaningful*
  benefit is far less unanimous than *any* benefit.
- EB shrinkage pulls observed effects 10–30 % toward the null. Marginal
  trials (DECAF, EMPA-REG) shrink most.
- The Design-informed prior re-creates the textbook DECAF "twist":
  the data and the protocol-prior point in opposite directions, so the
  posterior median is dragged materially closer to 1 and the calibration
  metrics flag the prior–data conflict.

## Methods

The pipeline follows four stages, all driven by a single curated
cohort CSV (`data/rct_cohort.csv`). The full source — including the
rehydrated EB mixture and `brms` model code — lives in `R/`:

1. **Cohort ingestion.** Trials are entered with reported effect (HR, RR),
   95% CI, raw arm counts where available, design power, MCID (the effect
   the trial was powered to detect), and an optional `design_hr` column for
   trials whose pre-specified direction differs from the observed effect
   direction (DECAF). Provenance for each row is documented in
   `docs/cohort.md`.
2. **Per-study frequentist layer.** From the reported HR and CI we recover
   `(log-HR, SE_logHR)` analytically, then derive `z`, `p`, and *two*
   Shannon S-values: `S vs RR = 1` (against the null) and `S vs RR = MCID`
   (against the design-target effect). Type-S and Type-M errors are
   computed by Gelman–Carlin retrodesign simulation given each trial's
   reported power.
3. **Empirical-Bayes posterior.** The four-component van Zwet 2025
   mixture on the *signal-to-noise-ratio* (SNR) scale is updated
   conjugately within each component and re-weighted by component
   likelihood, exactly matching `eb2.qmd`. We retain the full mixture
   posterior — not a Normal approximation — and propagate it to log-HR
   draws via `mu = SNR · SE_logHR`. Per-study mixture decomposition
   (component weights, conditional means, shrinkage factors `B_k`,
   pooled `B̄`) is reported.
4. **Bayesian re-analysis under five priors.** A `brms` normal
   pseudo-likelihood on `(logHR, SE_logHR)` is fit under each of:
   *Weak* (`N(0, 10)`), *Skeptical* (`N(0, 0.354)` — half on RR ≤ 0.5),
   *Enthusiastic* (`N(log MCID, 0.354)`), *EB* (van Zwet mixture, used
   directly via SNR draws — not re-fit through brms), and
   *Design-informed* (`N(log design_hr, SE_logHR)`). When no Stan
   backend is available, the four Normal priors fall back to the
   closed-form Normal–Normal posterior (mathematically identical to the
   `brms` fit for this likelihood); the EB row is always taken from the
   mixture draws.

The analytic outputs per study are: posterior median HR, 95% CrI,
`P(HR < 1)`, `P(HR ≤ MCID)`, plus Type-S, median Type-M, posterior
coverage, and same-sign replication probability. Cross-study summaries
are a forest plot of posterior medians by prior, a heatmap of
`P(HR ≤ MCID)`, and a shrinkage scatter (`|z|` vs posterior `|SNR|`).

## Cohort

```{r}
#| label: data
cohort <- readr::read_csv("data/rct_cohort.csv", show_col_types = FALSE)
knitr::kable(cohort |>
               dplyr::select(study, journal, year, endpoint, effect_type,
                             hr, ci_lo, ci_hi, n_t, n_c, power, mcid_rr))
```

```{r}
#| label: derived-quantities
# Derive logHR, SE, z, p, and TWO Shannon S-values per HR-scale row.
#
# Greenland's S-value (surprisal) is S = -log2(p).  It measures the
# information, in bits, against the hypothesis that generated p.
# S = k bits means the data are as surprising as k heads in a row from
# a fair coin under H0.
#
# We report two S-values per study:
#   * S_null : evidence against H0: RR = 1      (the textbook null)
#   * S_mcid : evidence against H_MCID: RR = mcid_rr
#               (evidence that the true effect is NOT at the
#                clinically-meaningful threshold the trial was powered for)
# Reading guide:
#   - Large S_null AND small S_mcid  -> strong evidence of benefit, but
#                                       the size is plausibly around MCID.
#   - Large S_null AND large S_mcid  -> strong benefit AND larger than MCID.
#   - Small S_null                    -> weak evidence of any effect.
cohort_hr <- cohort |>
  dplyr::filter(effect_type == "HR") |>
  dplyr::mutate(
    logHR      = log(hr),
    se_logHR   = (log(ci_hi) - log(ci_lo)) / (2 * 1.96),
    z_obs      = logHR / se_logHR,                       # z against RR = 1
    p_two      = 2 * pnorm(-abs(z_obs)),                 # p against RR = 1
    s_null     = s_value(pmax(p_two, .Machine$double.eps)),
    z_mcid     = (logHR - log(mcid_rr)) / se_logHR,      # z against RR = mcid
    p_mcid     = 2 * pnorm(-abs(z_mcid)),
    s_mcid     = s_value(pmax(p_mcid, .Machine$double.eps))
  )

# Keep a back-compat alias column for downstream code.
cohort_hr$s_value <- cohort_hr$s_null

knitr::kable(
  cohort_hr |>
    dplyr::transmute(
      study,
      logHR    = round(logHR, 3),
      se_logHR = round(se_logHR, 3),
      z        = round(z_obs, 2),
      p        = signif(p_two, 2),
      `S vs RR=1`    = round(s_null, 2),
      `S vs RR=MCID` = round(s_mcid, 2)
    ),
  caption = paste0(
    "Observed-effect frequentist summary. Both S-values are Shannon ",
    "surprisal values (bits). Larger = more evidence against the ",
    "reference RR. S = -log2(p)."
  )
)
```

### Interpreting the two S-values
Greenland's **S-value** (surprisal value) re-expresses a p-value as
bits of information against the tested hypothesis: `S = -log2(p)`.  Put
concretely, an S of 4.3 bits is *as surprising* as 4 or 5 heads in a row
from a fair coin under H0.  Reporting S together with the p-value gives
readers an intuitive sense of evidential strength without the 0.05 cliff.

- **`S vs RR = 1`** — the standard interpretation: evidence against no
  effect.  DECAF's 6.66 bits is the equivalent of 6–7 consecutive coin
  flips landing the same way under the null; PARADIGM-HF's 20 bits is
  roughly one-in-a-million territory.
- **`S vs RR = MCID`** — evidence that the true effect is *not* at the
  clinically-meaningful threshold the trial was powered for.  A small
  value here (close to zero) means the observed point estimate is
  statistically compatible with the MCID (which is the more interesting
  finding for borderline-significant trials like DECAF — strong evidence
  *against the null*, but negligible evidence against the MCID).

## Per-study analysis

```{r}
#| label: per-study
#| cache: true
run_one <- function(row) {
  study <- row$study
  hr    <- row$hr
  lo    <- row$ci_lo
  hi    <- row$ci_hi
  se    <- row$se_logHR
  logHR <- row$logHR
  z     <- row$z_obs
  mcid  <- row$mcid_rr
  # design_hr = mcid_rr unless the cohort row explicitly encodes a
  # pre-registered direction opposite to what was observed (e.g. DECAF).
  design_hr <- if (!is.null(row$design_hr) && is.finite(row$design_hr) &&
                   row$design_hr > 0) row$design_hr else mcid
  power <- row$power

  # --- Type M / Type S at the trial's design (unit-SNR scale) ---
  tms <- type_m_s(power = power, target_effect = 1, n_sim = 50000)

  # --- EB posterior on SNR scale (van Zwet 4-component mixture) ---
  # `eb$snr` are draws from the FULL mixture posterior on the SNR scale,
  # exactly as `eb2.qmd` produces them. We use these draws directly
  # (multiplied by SE) for the EB row so we faithfully propagate the
  # mixture's posterior, instead of squashing it into a Normal
  # approximation and re-running it through another Normal-Normal
  # update with the same data.
  eb <- eb_posterior_snr(z, n = N_EB)
  eb_mu_log <- eb$post_mean * se        # used by build_priors() only as a
  eb_sd_log <- eb$post_sd   * se        # Normal proxy on log-HR for brms.
  snr_post_mean <- eb$post_mean

  # Direct mixture-EB posterior on log-HR (the `eb2.qmd` computation):
  eb_mu_draws  <- eb$snr * se
  eb_RR_draws  <- exp(eb_mu_draws)
  eb_draws_tbl <- tibble::tibble(prior = "EB",
                                 mu    = eb_mu_draws,
                                 RR    = eb_RR_draws)

  # --- Five MCMC priors via brms ---
  priors <- build_priors(
    se_logHR  = se,
    mcid_rr   = mcid,
    eb_mu_log = eb_mu_log,
    eb_sd_log = eb_sd_log,
    design_hr = design_hr
  )
  df_fit <- tibble::tibble(effect = logHR, se = se)

  draws <- if (isTRUE(HAVE_STAN)) {
    tryCatch(
      suppressMessages(suppressWarnings(
        fit_brms_study(df_fit, priors, mcid = mcid,
                       chains = N_CHAINS, iter = N_ITER)
      )),
      error = function(e) NULL
    )
  } else {
    NULL
  }

  if (is.null(draws)) {
    # Analytic Normal-Normal fallback for the four Normal priors.
    analytic_post <- function(mu0, tau0) {
      prec_post <- 1 / tau0^2 + 1 / se^2
      mu_post   <- (mu0 / tau0^2 + logHR / se^2) / prec_post
      sd_post   <- sqrt(1 / prec_post)
      mu_draws  <- rnorm(4000, mu_post, sd_post)
      tibble::tibble(mu = mu_draws, RR = exp(mu_draws))
    }
    draws <- dplyr::bind_rows(
      analytic_post(0,              10)    |> dplyr::mutate(prior = "Weak"),
      analytic_post(0,              0.354) |> dplyr::mutate(prior = "Skeptical"),
      analytic_post(log(mcid),      0.354) |> dplyr::mutate(prior = "Enthusiastic"),
      analytic_post(log(design_hr), se)    |> dplyr::mutate(prior = "DesignInformed")
    )
  } else {
    # If brms produced draws under all five priors, drop its EB row and
    # replace it with the direct mixture-EB draws (more faithful to the
    # van Zwet posterior).
    draws <- draws |> dplyr::filter(prior != "EB")
  }
  draws <- dplyr::bind_rows(draws, eb_draws_tbl)

  summary_tbl <- summarise_posterior(draws, mcid = mcid) |>
    dplyr::mutate(study = study, .before = 1)

  list(
    study      = study,
    tms        = tms,
    eb         = list(snr_post_mean = snr_post_mean,
                      snr_draws     = eb$snr,
                      components    = eb$components,
                      B_bar         = eb$B_bar,
                      eb_mu_log     = eb_mu_log,
                      eb_sd_log     = eb_sd_log),
    draws      = draws,
    summary_tbl = summary_tbl
  )
}

results <- purrr::map(split(cohort_hr, seq_len(nrow(cohort_hr))), run_one)
names(results) <- cohort_hr$study
```

```{r}
#| label: study-plots
#| fig-width: 8
#| fig-height: 4
for (r in results) {
  mcid_rr <- cohort_hr$mcid_rr[cohort_hr$study == r$study]
  print(plot_study_densities(r$draws, mcid_rr, study = r$study))
}
```

```{r}
#| label: per-study-summary
summary_all <- dplyr::bind_rows(lapply(results, `[[`, "summary_tbl"))
knitr::kable(summary_all |> dplyr::mutate(
  median_RR = round(median_RR, 3),
  lo95 = round(lo95, 3),
  hi95 = round(hi95, 3),
  P_RR_lt_1 = round(P_RR_lt_1, 3),
  P_RR_le_mcid = round(P_RR_le_mcid, 3)
))

# Per-study frequentist-calibration table:
# Type-S / Type-M from the retrodesign simulation PLUS two Shannon
# S-values computed from the two-sided p against different reference
# RRs (null and MCID).
tms_tbl <- tibble::tibble(
  study   = vapply(results, `[[`, character(1), "study"),
  type_s  = vapply(results, function(r) r$tms$type_s, numeric(1)),
  type_m  = vapply(results, function(r) r$tms$type_m, numeric(1))
) |>
  dplyr::left_join(
    cohort_hr |> dplyr::select(study, z = z_obs, p = p_two,
                               s_null, s_mcid, mcid_rr),
    by = "study"
  )
knitr::kable(tms_tbl |>
               dplyr::mutate(type_s         = round(type_s, 3),
                             type_m         = round(type_m, 2),
                             `z`            = round(z, 2),
                             `p`            = signif(p, 2),
                             `S vs RR=1`    = round(s_null, 2),
                             `S vs RR=MCID` = round(s_mcid, 2),
                             `MCID`         = mcid_rr) |>
               dplyr::select(study, type_s, type_m, z, p,
                             `S vs RR=1`, `MCID`, `S vs RR=MCID`),
             caption = paste0(
               "Frequentist calibration: Type-S, Type-M exaggeration, ",
               "observed z, p, and two Shannon S-values. ",
               "A large 'S vs RR=1' with a small 'S vs RR=MCID' means ",
               "the data strongly reject the null but are compatible ",
               "with an effect at the design-MCID magnitude."))
```

## Empirical-Bayes deep-dive (per `eb2.qmd`)
### Cochrane mixture component weights, by study
For each study, the table below shows how the **posterior weight** is
redistributed across the four van Zwet components after seeing the
observed z. Smaller-SD components correspond to more aggressive shrinkage
(`B_k` close to 0); larger-SD components leave the data nearly untouched
(`B_k` close to 1). The `B_bar` column is the overall posterior shrinkage
factor (a study-level summary of how much the EB prior moves the raw
estimate toward 0).

```{r}
#| label: eb-components
eb_comp_tbl <- dplyr::bind_rows(lapply(names(results), function(s) {
  comp <- results[[s]]$eb$components
  comp$component <- seq_len(nrow(comp))
  comp$study     <- s
  comp$B_bar     <- results[[s]]$eb$B_bar
  comp
}))
knitr::kable(
  eb_comp_tbl |>
    dplyr::select(study, component, weight, mean, sd, B, B_bar) |>
    dplyr::mutate(weight = round(weight, 3),
                  mean   = round(mean,   3),
                  sd     = round(sd,     3),
                  B      = round(B,      3),
                  B_bar  = round(B_bar,  3)),
  caption = paste0(
    "Per-study posterior decomposition under the van Zwet mixture: ",
    "component-wise weight, conditional mean (B_k * z), SD, shrinkage ",
    "factor B_k, and overall pooled shrinkage B_bar."
  )
)
```

### Calibration metrics (Type-S / Type-M / Coverage / Replication)
These are the same four calibration metrics computed by `eb2.qmd`'s
`compute_metrics()` function, evaluated for every (study, prior)
combination using posterior draws on the SNR scale
(`SNR_draws = mu_draws / SE`):

- **Type-S** = `P(sign(SNR_post) != sign(z_obs))` - probability the true
  effect has the opposite sign to the observed z.
- **Type-M (median)** = `median(|z_obs / SNR_post|)` - typical exaggeration
  of |z| relative to the true effect size on the SNR scale.
- **Coverage** = `P(z_obs - 1.96 < SNR_post < z_obs + 1.96)` - the
  fraction of posterior SNR mass inside a same-size CI around the
  observed z (high values flag a posterior nearly identical to the data;
  low values flag aggressive shrinkage).
- **Replication** = `P(|SNR_post + N(0,1)| > 1.96  AND  same sign as z_obs)`
  - Bayesian probability that a same-size replication trial would land
  outside +/- 1.96 *with the original sign*.

```{r}
#| label: eb-calibration-metrics
compute_eb_metrics <- function(z_obs, snr_draws) {
  if (!length(snr_draws) || all(!is.finite(snr_draws))) {
    return(data.frame(type_s     = NA_real_,
                      type_m_med = NA_real_,
                      coverage   = NA_real_,
                      replicate  = NA_real_))
  }
  infl  <- abs(z_obs / snr_draws)
  s_err <- mean(sign(z_obs) != sign(snr_draws), na.rm = TRUE)
  cover <- mean(snr_draws > (z_obs - 1.96) &
                snr_draws < (z_obs + 1.96), na.rm = TRUE)
  z2    <- snr_draws + rnorm(length(snr_draws))
  repl  <- mean(abs(z2) > 1.96 & sign(z2) == sign(z_obs), na.rm = TRUE)
  data.frame(type_s     = s_err,
             type_m_med = stats::median(infl, na.rm = TRUE),
             coverage   = cover,
             replicate  = repl)
}

set.seed(2026)
calib_tbl <- dplyr::bind_rows(lapply(names(results), function(s) {
  z   <- cohort_hr$z_obs[cohort_hr$study == s]
  se  <- cohort_hr$se_logHR[cohort_hr$study == s]
  drs <- results[[s]]$draws
  prs <- split(drs, drs$prior)
  rows <- lapply(names(prs), function(p) {
    snr <- prs[[p]]$mu / se     # log-HR draws -> SNR draws
    cbind(study = s, prior = p, compute_eb_metrics(z, snr))
  })
  do.call(rbind, rows)
}))

knitr::kable(
  calib_tbl |>
    dplyr::mutate(type_s     = round(type_s,     3),
                  type_m_med = round(type_m_med, 2),
                  coverage   = round(coverage,   3),
                  replicate  = round(replicate,  3)) |>
    dplyr::arrange(study, factor(prior, levels = c(
      "Weak", "Skeptical", "Enthusiastic", "EB", "DesignInformed"))),
  caption = paste0(
    "Posterior-based calibration metrics by study and prior ",
    "(matches eb2.qmd `compute_metrics()`). Type-S, Type-M-median, ",
    "Coverage, and same-sign Replication probability."
  )
)
```

Reading the table: under **EB** every study has Type-S near 0 and
Type-M-median near 1 (the mixture prior aggressively shrinks any
posterior that would imply a sign error or extreme exaggeration). Under
**DesignInformed**, DECAF stands out: a Type-S that is no longer 0
(the prior pulls a non-trivial fraction of posterior mass to the
opposite side of the observed z), and a coverage that drops sharply -
the data and the design-target prior simply disagree.

## Cross-study summary

```{r}
#| label: forest
#| fig-width: 9
#| fig-height: 5
plot_forest(summary_all)
ggsave(file.path(RESULTS_DIR, "forest.png"), width = 9, height = 5, dpi = 120)
```

```{r}
#| label: heatmap
#| fig-width: 8
#| fig-height: 5
plot_heatmap_mcid(summary_all)
ggsave(file.path(RESULTS_DIR, "heatmap_mcid.png"), width = 8, height = 5, dpi = 120)
```

```{r}
#| label: shrinkage-scatter
#| fig-width: 7
#| fig-height: 5
shrink_df <- tibble::tibble(
  study        = cohort_hr$study,
  abs_z        = abs(cohort_hr$z_obs),
  abs_snr_post = vapply(results,
                        function(r) abs(r$eb$snr_post_mean),
                        numeric(1))
)
plot_shrinkage_scatter(shrink_df)
ggsave(file.path(RESULTS_DIR, "shrinkage_scatter.png"), width = 7, height = 5, dpi = 120)

# Persist a flat summary CSV for run_pipeline.R consumers.
readr::write_csv(
  dplyr::left_join(summary_all, tms_tbl, by = "study") |>
    dplyr::left_join(cohort_hr |>
                       dplyr::select(study, hr, ci_lo, ci_hi, z_obs, p_two,
                                     s_value, power, mcid_rr), by = "study"),
  file.path(RESULTS_DIR, "pipeline_summary.csv")
)
```

## DECAF worked example (validation)

DECAF randomised 200 post-cardioversion AF patients (1:1) to **caffeinated
coffee (~1 cup/day)** versus **caffeine abstinence** [@RN1]. The published
result — HR = 0.61 (95% CI 0.42 - 0.89, p = 0.01) — is for *caffeinated
vs abstinence*: **caffeinated had 39% lower AF recurrence**.

The trial was powered in the **opposite direction**. Caffeine has
historically been viewed as proarrhythmic, so the protocol was designed
to detect a **41% reduction in AF with *abstinence*** (decaf). In the
caff/abst frame used by the published HR, that design target corresponds
to a *hazard ratio greater than 1*: `design_hr = 1 / 0.59 ≈ 1.695`
(i.e. the trial expected caffeinated to INCREASE AF vs abstinence).

So DECAF gives us two quantities on opposite sides of 1:

- **Observed effect:** HR = 0.61 (caffeinated favours) — log-HR ≈ -0.49
- **Design target:**   HR = 1.695 (abstinence favours, pre-specified) — log-HR ≈ +0.53

The **Design-informed prior** is therefore built around `log(design_hr)`,
not `log(mcid_rr)`, and will **conflict** with the observed data. This is
the point of the DECAF example in `eb2.qmd`.

The **clinically meaningful threshold** in the observed direction is still
`mcid_rr = 0.59` (a 41% reduction in the caffeinated arm would match the
magnitude the trial was powered for, just with the arms swapped). That is
what `P(HR <= MCID)` reports below.

```{r}
#| label: decaf
decaf <- cohort_hr |> dplyr::filter(study == "DECAF")
stopifnot(nrow(decaf) == 1)
r_decaf <- results[["DECAF"]]

# Observed summary + frequentist calibration (both S-values).
decaf_obs <- dplyr::transmute(
  decaf,
  `HR (observed)`  = hr,
  `95% CI`         = sprintf("[%.2f, %.2f]", ci_lo, ci_hi),
  `log HR`         = round(logHR, 3),
  `SE[log HR]`     = round(se_logHR, 3),
  z                = round(z_obs, 2),
  p                = signif(p_two, 2),
  `S vs RR=1`      = round(s_null, 2),
  `S vs RR=MCID`   = round(s_mcid, 2),
  `MCID (RR)`      = mcid_rr,
  `Design HR`      = design_hr
)
knitr::kable(decaf_obs, caption = "DECAF: frequentist observables, both Shannon S-values, and direction markers")

# Posterior summaries under all five priors.
decaf_post <- r_decaf$summary_tbl |>
  dplyr::mutate(dplyr::across(median_RR:P_RR_le_mcid, ~round(.x, 3)))
knitr::kable(decaf_post,
             caption = "DECAF posterior HR summary across the five priors")

# Key quantities for the narrative.
eb_row <- decaf_post |> dplyr::filter(prior == "EB")
di_row <- decaf_post |> dplyr::filter(prior == "DesignInformed")
eb_rr_med <- eb_row$median_RR
di_rr_med <- di_row$median_RR

cat(sprintf("DECAF EB posterior HR median           : %.3f (95%% CrI %.3f - %.3f)\n",
            eb_rr_med, eb_row$lo95, eb_row$hi95))
cat(sprintf("DECAF Design-informed posterior median : %.3f (95%% CrI %.3f - %.3f)\n",
            di_rr_med, di_row$lo95, di_row$hi95))

# Sanity check: EB posterior HR median should lie in [0.65, 0.90].
stopifnot(eb_rr_med >= 0.65, eb_rr_med <= 0.90)

# Sanity check: the Design-informed prior, being centred on the opposite
# side of 1 to the observed data, must produce a posterior materially
# closer to 1 (and above the EB median).
stopifnot(di_rr_med > eb_rr_med)
```

Interpretation:

- Under **EB** shrinkage (calibrated from ~23,551 Cochrane trials), the
  posterior HR shrinks modestly from the raw 0.61 towards 1, landing near
  `r round(eb_rr_med, 3)` with 95% CrI `r round(eb_row$lo95, 2)` to
  `r round(eb_row$hi95, 2)`. `P(HR < 1)` is essentially 1.
- Under the **Design-informed** prior (centred on `HR = 1.695` per the
  trial protocol), the posterior is pulled *away* from the observed
  effect: median HR lands near `r round(di_rr_med, 3)`. This is the
  price of trusting a prior that turned out to contradict the data.
- `P(HR <= MCID = 0.59)` under EB is a modest `r eb_row$P_RR_le_mcid` —
  i.e., the probability that caffeinated produces at least a 41%
  reduction in AF recurrence (the magnitude the trial was powered for,
  just in the opposite arm) is around that value.
- **S-values tell the same story in bits.** `S vs RR=1` for DECAF is
  `r round(decaf$s_null, 2)` bits — strong evidence *against* no effect,
  as surprising as `r round(decaf$s_null, 0)` consecutive coin flips
  under H0. But `S vs RR=MCID` is only `r round(decaf$s_mcid, 2)` bits:
  the observed point estimate (HR 0.61) is **statistically compatible**
  with the MCID magnitude (HR 0.59) — there is essentially no evidence
  that the true effect differs from the clinically meaningful threshold.
  Together, the two S-values show the classic profile of a borderline
  significant small trial: clearly *something*, but not clearly *big*.

## Discussion

Across the curated cohort, three patterns emerge.

**"Some benefit" is robust; "clinically meaningful benefit" is not.**
`P(HR < 1)` is at or above 0.98 for every significant trial under every
prior. `P(HR ≤ MCID)`, in contrast, ranges from near-zero for marginal
trials (EMPA-REG ≈ 0.05; DECAF ≈ 0.22 under EB) to near-one for landmark
trials with effect sizes well past their design target (PARADIGM-HF,
RALES). Reporting only `P(HR < 1)` therefore exaggerates how confidently
a trial achieves the magnitude of effect it was sold on.

**EB shrinkage is study-strength dependent.** The pooled posterior
shrinkage factor `B̄` — which is what closes the gap between observed
and true effect under the van Zwet mixture — is close to 1 (minimal
shrinkage) for high-`|z|` trials and drops materially for borderline
trials. Operationally, this means EB *acts the most* exactly where it
is most needed: marginal results where Type-M exaggeration is highest.

**Design-informed priors can be misleading when the protocol gets it
wrong.** DECAF is the cleanest example: the protocol expected
abstinence to *reduce* AF recurrence, but the trial found that
caffeinated coffee did. Centering a prior on `log(design_hr) ≈ +0.53`
drags the posterior toward 1 and inflates Type-S, even though the
frequentist evidence against the null is comfortably > 6 bits. This is
a warning against treating the powered-effect direction as a default
prior when the literature on which the design rests was itself
uncertain.

These patterns reinforce the methodological recommendations of van Zwet
et al. and the "abandon statistical significance" literature: report
posterior probabilities at *clinically meaningful* thresholds, not just
at the null, and surface the Type-M / shrinkage diagnostics alongside
the headline estimate.

## Limitations

- The cohort is a *curated* convenience sample of landmark CV RCTs and
  is not the result of a systematic database search; relative effect
  sizes here are illustrative of the methodology rather than
  representative of the cardiovascular literature as a whole.
- The `brms` model is a normal pseudo-likelihood on summary
  `(logHR, SE)` rather than a re-analysis of patient-level data.
  Where individual-patient data are available, refitting with arm-level
  event counts (`brms::brm(events | trials(n) ~ arm, family = binomial())`)
  would be preferred.
- The van Zwet mixture is calibrated on the Cochrane Database of
  Systematic Reviews; its applicability to specialty cardiovascular
  endpoints (e.g. composite MACE) is plausible but not separately
  validated here.
- MCID values come from the original power calculations and inherit
  whatever optimism was embedded there; sensitivity to alternative
  MCID choices is implicit in the `S vs RR = MCID` column but not
  formally varied across the cohort in this version.
- The Stan probe in setup detects `cmdstanr`/`rstan` and silently falls
  back to the analytic Normal–Normal posterior if neither is available.
  Numerically the two paths are equivalent for this likelihood, but
  reproducing the exact `brms` chain diagnostics requires a working
  Stan toolchain.

## Reproducibility {.unnumbered}

```{r}
#| label: session-info
#| echo: false
sessionInfo()
```

```{r}
#| label: env-summary
#| echo: false
cat(sprintf("Stan backend available : %s\n", isTRUE(HAVE_STAN)))
cat(sprintf("MCMC chains            : %d\n", N_CHAINS))
cat(sprintf("MCMC iterations / chain: %d\n", N_ITER))
cat(sprintf("EB mixture draws       : %d\n", N_EB))
cat(sprintf("Cohort rows analysed   : %d\n", nrow(cohort_hr)))
cat(sprintf("Pipeline summary CSV   : %s\n",
            normalizePath(file.path(RESULTS_DIR, "pipeline_summary.csv"),
                          mustWork = FALSE)))
```

## References {.unnumbered}
